---
title: "Homework3"
author: "Robert Smigielski"
format: html
editor: visual
---

## Homework 3

**Loads libraries**

```{r}
library(tidyverse)
```

## 2.1

**Reads the dataset**

```{r}
rhyming_data <- read.csv("C:/Users/smigi/Desktop/PS947/rhyming.csv")
```

**Displays the first few rows of the dataset to check its structure**

```{r}
head(rhyming_data)
```

**`unite()` used to merge 'image_1' and 'image_2', into a single 'stimulus' column**

**Underscore was used to separate the two names for clarity**

**`remove = TRUE` removes the original 'image_1' and 'image_2' columns after merging**

```{r}
rhyming_data <- rhyming_data %>% 
  unite(col = "stimulus", image_1, image_2, sep = "_", remove = TRUE)
```

**Converts 'type' to a factor variable to make modeling and plotting easier later**

```{r}
rhyming_data$type <- as.factor(rhyming_data$type)
```

**Converts 'high_low_verbal' to a factor as well, as it is categorical data**

```{r}
rhyming_data$high_low_verbal <- as.factor(rhyming_data$high_low_verbal)
```

**Checks for missing values for possible data cleaning**

```{r}
sum(is.na(rhyming_data)) # Returns total count of missing values
```

**Displays structure of the cleaned dataset to check changes**

```{r}
str(rhyming_data)
```

**Saves the cleaned dataset for further analysis**

```{r}
write.csv(rhyming_data, "tidy_rhyming.csv", row.names = FALSE)
```
