---
title: "Homework3"
author: "Robert Smigielski"
format: html
editor: visual
---

# Homework 3

**Loads libraries**

```{r}
library(tidyverse)
library(ggplot2)
```

## 2.1

**Reads the dataset**

```{r}
rhyming_data <- read.csv("C:/Users/smigi/Desktop/PS947/rhyming.csv")
```

**Displays the first few rows of the dataset to check its structure**

```{r}
head(rhyming_data)
```

**`unite()` used to merge 'image_1' and 'image_2', into a single 'stimulus' column**

**Underscore was used to separate the two names for clarity**

**`remove = TRUE` removes the original 'image_1' and 'image_2' columns after merging**

```{r}
rhyming_data <- rhyming_data %>% 
  unite(col = "stimulus", image_1, image_2, sep = "_", remove = TRUE)
```

**Converts 'type' to a factor variable to make modeling and plotting easier later**

```{r}
rhyming_data$type <- as.factor(rhyming_data$type)
```

**Converts 'high_low_verbal' to a factor as well, as it is categorical data**

```{r}
rhyming_data$high_low_verbal <- as.factor(rhyming_data$high_low_verbal)
```

**Checks for missing values**

```{r}
sum(is.na(rhyming_data)) # Returns total count of missing values
```

**Displays structure of the dataset to check changes**

```{r}
str(rhyming_data)
```

## 2.2

**Plots the data to show how stimulus type and the level of inner voice affect reaction times**

-   **`aes()` defines the aesthetics for the plot:**

-   **`x = type` maps the stimulus type to the x-axis**

-   **`y = rt` maps the reaction time (ms) to the y-axis**

-   **`fill = high_low_verbal` maps the inner voice level (high or low) to the fill colour of the boxes.**

```{r}
plot <- ggplot(rhyming_data, aes(x = type, y = rt, fill = high_low_verbal)) + 
  geom_boxplot() + # Creates boxplots for each combination of `type` and `high_low_verbal`
  scale_fill_manual(values = c("high_verbal" = "skyblue", "low_verbal" = "salmon")) + # Fills boxplot with custom colors
  labs(title = "Effect of Stimulus Type and Inner Voice Level on Reaction Times",
       x = "Stimulus Type",
       y = "Reaction Time (ms)",
       fill = "Inner Voice Level") +
  theme_minimal() # Applies a minimal theme to the plot for a clean look
```

**Displays the plot**

```{r}
print(plot)
```

**Saves the plot to a file with high resolution**

```{r}
ggsave("reaction_time_plot.png", plot = plot, dpi = 300)
```
